Demo arXiv template

### 2023-02-11 paper/paper.qmd
add example on calculating indexes + two captions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(width = 90)
library(tidyverse)
library(latex2exp)
library(patchwork)
library(indri)
library(GGally)
```

@@

```{r eval = FALSE}
DATA %>% 
  init(id = ..., time = ...) %>% 
  compute_indexes(
    spei = idx_spei(...),
    spi = idx_spei(...)
  )
```

@@

For SPI and SPEI, the wrapper functions also allow for multiple inputs of scales and distribution parameters (`.scale` and `.dist`) to allow computation on multiple time scale and distribution fitting. The following code compute the two indexes (SPI and SPEI) for different time scales (6, 12, 18, 24, 30, 36) and different distributions (Pearson III, log-logistic, lognormal, and general extreme value distributions) for [NUMBER] weather stations in Eastern Australia shown in @fig-station-map: (TODO: should talk about the data source)

```{r eval = FALSE}
scale <- c(6, 12, 18, 24, 30, 36)
dist_spei <- c(gev(), loglogistic(), pearsonIII(), lognormal())
dist_spi <- c(gamma(), weibull())
dt %>% 
  compute_indexes(
    spei = idx_spei(
      .pet_method = "thornthwaite", .tavg = tavg, .lat = lat, 
      .scale = scale, .dist = dist_spei),
    spi = idx_spi(.scale = scale,  .dist = dist_spi)
  )
```

```{r fig-station-map}
#| fig-cap: "..."
#| echo: false

map_dt <- ozmaps::abs_ste %>%
  filter(NAME %in% c("New South Wales", "Victoria", "Queensland")) %>%
  rmapshaper::ms_simplify(keep = 0.01)

map_dt %>%
  ggplot() +
  geom_sf(fill = "transparent") +
  theme_void() 
  #geom_point(data = , aes(x = long, y = lat))
```

-   a basic workflow and congruence with results in the `SPEI` pkg
-   allow multiple distribution fit
-   allow bootstrap uncertainty

easier to make comparison
pick a publication, propose a new index that has this benefit but also correlate highly with the big name indexes

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(width = 90)
library(tidyverse)
library(latex2exp)
library(patchwork)
library(indri)
library(lmomco)
library(lubridate)
library(SPEI)
library(tsibble)
library(GGally)
library(patchwork)
```

@@

```{r eval = FALSE, echo = TRUE}
DATA %>% 
  init(id = ..., time = ...) %>% 
  compute_indexes(
    spei = idx_spei(...),
    spi = idx_spei(...)
  )
```

@@

For SPI and SPEI, the wrapper functions also allow for multiple inputs of scales and distribution parameters (`.scale` and `.dist`) to allow computation on multiple time scale and distribution fitting.

The climate data is sourced from Global Historical Climatology Network Daily (GHCND).

stations in Queensland, Australia. 
With precipitation, maximum and minimum temperature available since January 1990.

This gives `r length(unique(queensland$id))` stations.

The following code compute the two indexes (SPI and SPEI) at different time scales (6, 12, 24, 36) and different distributions (log-logistic and general extreme value distributions for SPEI and gamma for SPI).

```{r cache=TRUE, echo = TRUE}
.scale <- c(6, 12, 24, 36)
(idx <- queensland %>%
  init(id = id, time = ym) %>%
  compute_indexes(
    spei = idx_spei(
      .pet_method = "thornthwaite", .tavg = tavg, .lat = lat,
      .scale = .scale, .dist = c(gev(), loglogistic())),
    spi = idx_spi(.scale = .scale)
  ))
```

The output from `compute_indexes()` contains index values and associated parameter in a long tibble. Analysis and visualisation can then be conducted to understand the effect of scale and distribution on the index value or explore the spatial and temporal pattern of wet/dry condition as in @fig-compute-spatial and @fig-compute-temporal.

```{r}
library(sf)
qld_lga <- absmapsdata::lga2018 %>%
  filter(state_name_2016 == "Queensland") %>%
  rmapshaper::ms_simplify(keep = 0.3) %>%
  mutate(lga_name_2018 = str_replace(lga_name_2018, " \\([S|R|C|]\\)", ""),
         lga_name_2018 = str_replace(lga_name_2018, " \\(Qld\\)", ""))
```

<!-- maybe combine the two figures: Exploratory data analysis on the computed indexes: ??? -->

```{r fig-compute-spatial}
#| fig-cap: "Spatial distribution of Standardized Precipitation Index (SPI-12) in Queensland, Australia during two major flood and drought events: 2010/11 and 2019/20. The map shows a continous wet period during the 2010/11 flood period and a mitigated drought situation, after its worst in 2019 December and 2020 Janurary, likely due to the increased rainfall in February from the meteorological record."
#| fig-width: 12
#| fig-height: 5
queensland_map <- ozmaps::abs_ste %>% filter(NAME == "Queensland") %>% 
  rmapshaper::ms_simplify(keep = 0.02)
queensland_map %>%
  ggplot() +
  geom_sf(fill = "transparent", color = "grey50", linewidth = 0.4) +
  geom_point(
    data = idx %>%
      filter(.idx == "spi", .scale == 12) %>%
      filter(((year(ym) %in% c(2010, 2019) & month(ym) >= 10) )|
               ((year(ym) %in% c(2011, 2020) & month(ym) <= 3) )), 
    aes(x = long, y = lat, color = .index)) +
  scale_color_distiller(palette = "BrBG", direction = 1, name = "Index", ) +
  theme_void() +
  facet_wrap(vars(ym), ncol = 6)
```

```{r fig-compute-temporal}
#| fig-cap: 'Time series plot of Standardized Precipitation-Evapotranspiration Index (SPEI) at the Texas post office station (highlighted by a diamond shape in panel a). The SPEI is calculated at four time scales (6, 12, 24, and 36 months) and fitted with two distributions (Log Logistic and GEV). The dashed line at -2 represents the class "extreme drought" by the SPEI. A larger time scale gives a smoother index series, while also takes longer to recover from an extreme situation as seen in the  2019/20 drought period. The SPEI values from two distribution fits mostly agree, while GEV can results in more extreme values, i.e. in 1998 and 2020.'
#| fig-width: 12
#| fig-height: 6
texas <- idx %>%
    filter(name == "TEXAS POST OFFICE", .idx == "spei") %>% distinct(long, lat)
p1 <- queensland_map %>%
  ggplot() +
  geom_sf(fill = "transparent") +
  geom_point(data = queensland %>%
               distinct(id, long, lat, name),
             aes(x = long, y = lat)) +
  geom_point(data = texas, aes(x = long, y = lat),
             color = "orange", shape = 18, fill = "orange", size = 4) +  
  theme_void()

p2 <- idx %>%
  filter(name == "TEXAS POST OFFICE", .idx == "spei") %>%
  mutate(.dist = ifelse(.dist == "pargev", "GEV", "Log Logistic")) %>%
  rename(scale = .scale) %>% 
  ggplot(aes(x = ym, y = .index, color = .dist, group = .dist)) +
  geom_line() +
  theme_benchmark() +
  facet_wrap(vars(scale), labeller = label_both, ncol = 1) +
  scale_x_yearmonth(breaks = "2 year", date_labels = "%Y") +
  scale_color_brewer(palette = "Dark2", name = "Distribution") + 
  xlab("Year") + 
  ylab("Index")

(p1 | p2)  + 
  patchwork::plot_annotation(tag_levels = "a") + 
  patchwork::plot_layout(guides = "collect")  &
  theme(legend.position = "bottom")

### 2023-02-11 paper/paper.tex
add example on calculating indexes + two captions

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden]}{\end{tcolorbox}}\fi

@@

@@

For SPI and SPEI, the wrapper functions also allow for multiple inputs
of scales and distribution parameters (\texttt{.scale} and
\texttt{.dist}) to allow computation on multiple time scale and
distribution fitting. The following code compute the two indexes (SPI
and SPEI) for different time scales (6, 12, 18, 24, 30, 36) and
different distributions (Pearson III, log-logistic, lognormal, and
general extreme value distributions) for {[}NUMBER{]} weather stations
in Eastern Australia shown in Figure~\ref{fig-station-map}: (TODO:
should talk about the data source)

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-station-map-1.pdf}

@@

\caption{\label{fig-station-map}\ldots{}}

@@

\begin{itemize}
\tightlist
\item
  a basic workflow and congruence with results in the \texttt{SPEI} pkg
\item
  allow multiple distribution fit
\item
  allow bootstrap uncertainty
\end{itemize}

easier to make comparison pick a publication, propose a new index that
has this benefit but also correlate highly with the big name indexes

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners, frame hidden, enhanced, interior hidden]}{\end{tcolorbox}}\fi

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{init}\NormalTok{(}\AttributeTok{id =}\NormalTok{ ..., }\AttributeTok{time =}\NormalTok{ ...) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{compute\_indexes}\NormalTok{(}
    \AttributeTok{spei =} \FunctionTok{idx\_spei}\NormalTok{(...),}
    \AttributeTok{spi =} \FunctionTok{idx\_spei}\NormalTok{(...)}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

@@

For SPI and SPEI, the wrapper functions also allow for multiple inputs
of scales and distribution parameters (\texttt{.scale} and
\texttt{.dist}) to allow computation on multiple time scale and
distribution fitting.

The climate data is sourced from Global Historical Climatology Network
Daily (GHCND).

stations in Queensland, Australia. With precipitation, maximum and
minimum temperature available since January 1990.

This gives 29 stations.

The following code compute the two indexes (SPI and SPEI) at different
time scales (6, 12, 24, 36) and different distributions (log-logistic
and general extreme value distributions for SPEI and gamma for SPI).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{.scale }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{12}\NormalTok{, }\DecValTok{24}\NormalTok{, }\DecValTok{36}\NormalTok{)}
\NormalTok{(idx }\OtherTok{\textless{}{-}}\NormalTok{ queensland }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{init}\NormalTok{(}\AttributeTok{id =}\NormalTok{ id, }\AttributeTok{time =}\NormalTok{ ym) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{compute\_indexes}\NormalTok{(}
    \AttributeTok{spei =} \FunctionTok{idx\_spei}\NormalTok{(}
      \AttributeTok{.pet\_method =} \StringTok{"thornthwaite"}\NormalTok{, }\AttributeTok{.tavg =}\NormalTok{ tavg, }\AttributeTok{.lat =}\NormalTok{ lat,}
      \AttributeTok{.scale =}\NormalTok{ .scale, }\AttributeTok{.dist =} \FunctionTok{c}\NormalTok{(}\FunctionTok{gev}\NormalTok{(), }\FunctionTok{loglogistic}\NormalTok{())),}
    \AttributeTok{spi =} \FunctionTok{idx\_spi}\NormalTok{(}\AttributeTok{.scale =}\NormalTok{ .scale)}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 128,586 x 19
   .idx  .period id              ym  prcp  tmax  tmin  tavg  long   lat name    pet   diff
   <chr>   <dbl> <chr>        <mth> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <chr> <dbl>  <dbl>
 1 spei        6 ASN00029~ 1990 Jun   170  29.7  16.2  23.0  142. -15.5 KOWA~  67.8  102. 
 2 spei        6 ASN00029~ 1990 Jun   170  29.7  16.2  23.0  142. -15.5 KOWA~  67.8  102. 
 3 spei        6 ASN00029~ 1990 Jun     0  23.0  11.8  17.4  139. -20.7 MOUN~  30.8  -30.8
 4 spei        6 ASN00029~ 1990 Jun     0  23.0  11.8  17.4  139. -20.7 MOUN~  30.8  -30.8
 5 spei        6 ASN00031~ 1990 Jun   794  25.8  18.1  21.9  146. -16.9 CAIR~  66.3  728. 
 6 spei        6 ASN00031~ 1990 Jun   794  25.8  18.1  21.9  146. -16.9 CAIR~  66.3  728. 
 7 spei        6 ASN00031~ 1990 Jun   504  23.0  13.8  18.4  145. -17.1 WALK~  48.4  456. 
 8 spei        6 ASN00031~ 1990 Jun   504  23.0  13.8  18.4  145. -17.1 WALK~  48.4  456. 
 9 spei        6 ASN00032~ 1990 Jun  1970  23.9  16.4  20.2  146. -17.6 SOUT~  54.5 1915. 
10 spei        6 ASN00032~ 1990 Jun  1970  23.9  16.4  20.2  146. -17.6 SOUT~  54.5 1915. 
# ... with 128,576 more rows, and 6 more variables: .scale <dbl>, .agg <dbl>,
#   .method <chr>, .fitted <dbl>, .dist <chr>, .index <dbl>
\end{verbatim}

The output from \texttt{compute\_indexes()} contains index values and
associated parameter in a long tibble. Analysis and visualisation can
then be conducted to understand the effect of scale and distribution on
the index value or explore the spatial and temporal pattern of wet/dry
condition as in Figure~\ref{fig-compute-spatial} and
Figure~\ref{fig-compute-temporal}.

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-compute-spatial-1.pdf}

@@

\caption{\label{fig-compute-spatial}Spatial distribution of Standardized
Precipitation Index (SPI-12) in Queensland, Australia during two major
flood and drought events: 2010/11 and 2019/20. The map shows a continous
wet period during the 2010/11 flood period and a mitigated drought
situation, after its worst in 2019 December and 2020 Janurary, likely
due to the increased rainfall in February from the meteorological
record.}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-compute-temporal-1.pdf}

\caption{\label{fig-compute-temporal}Time series plot of Standardized
Precipitation-Evapotranspiration Index (SPEI) at the Texas post office
station (highlighted by a diamond shape in panel a). The SPEI is
calculated at four time scales (6, 12, 24, and 36 months) and fitted
with two distributions (Log Logistic and GEV). The dashed line at -2
represents the class ``extreme drought'' by the SPEI. A larger time
scale gives a smoother index series, while also takes longer to recover
from an extreme situation as seen in the 2019/20 drought period. The
SPEI values from two distribution fits mostly agree, while GEV can
results in more extreme values, i.e.~in 1998 and 2020.}
